results

### 2020-07-26 analysis/results.Rmd
tmp save

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "g"))
    } else{
      return(x)
    }
  }
)
```

@@

Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio including hybrid journals. Figure shows that the majority of journals supported the hybrid model. Overall, Elsevier published `r length(unique(els_jns$issn))`, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.

Elseviers journal portfolio changed over the period of observation. The number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals between 2015 and May 2020, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%. The year 2016 marked the highest number of journals. After that, the trend reversed. Since 2018, the number of journals stagnated. On 23 May 2020, Elsevier APC pricing list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.

@@

Following the approach of our earlier study [@Matthias_2019], we furthermore used APC pricing list information to determine the number of open access business model flips between 2015-2020. Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed. A total number of `r sum(flipped_jns_ind$n)` journals changed the business model under the ownership of Elsevier. Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n` changed to hybrid open access. Examples for such reverse flipped open access journals include the physical journal *Materials Today* and the life science journal *Meta Gene*.

@@

# oa
ind_per_year <- hybrid_df %>%
  filter(oa_archive == FALSE) %>%
  group_by(issued_year) %>%
  summarise(oa_articles = n()) %>%
  inner_join(els_yr, by = "issued_year") %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles * 100)
ind_summary_table <- ind_per_year %>% 
  pivot_longer(oa_articles:prop, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count) %>%
  mutate(indicator = factor(indicator, levels = c("journals", "articles", "oa_articles", "prop"))) %>%
  arrange(indicator) %>%
  mutate(indicator = recode(indicator,
                            oa_articles = "Hybrid OA Articles", 
                            articles = "All Articles", 
                            journals = "Hybrid journals with at least one OA article", 
                            prop = "Percentage"))
`Total` <- c(
  hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
  sum(ind_summary_table[2,2:6]),
  sum(ind_summary_table[3,2:6]),
  sum(ind_per_year$oa_articles) / sum(ind_per_year$articles) * 100)
bind_cols(ind_summary_table, Total = `Total`) %>%
  rename(` ` = indicator) %>%
  knitr::kable(digits = 1)
```

Table presents the high-level findings by year of publication. illustrating a growth

>>>>

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(formatC(x, big.mark = ",", format = "fg", digits = 2))
    } else{
      return(x)
    }
  }
)
```

@@

Using text-mined data from web-archived Elsevier APC pricing lists  [@Matthias_2020], we were able to determine the evolution of Elsevier's open access journal portfolio. Figure shows that the majority of journals supported the hybrid model. Overall, Elsevier published `r length(unique(els_jns$issn))` open access journals, of which `r els_jns %>%  filter(oa_model == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model.

Elsevier's journal portfolio changed over the period of observation. The number of hybrid journals grew by  `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$n) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$n)` journals between 2015 and May 2020, while its proportion relative to Elsevier's overall journal portfolio slightly decreased from `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% .$prop, 2) * 100`% to `r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% .$prop, 2) * 100`%. The year 2016 marked the highest number of journals. After that, the trend reversed. Since 2018, the number of journals stagnated. On 23 May 2020, Elsevier APC pricing list comprised `r length(unique(els_jns_df$issn))` journals, of which `r els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow()` supported the hybrid model, representing a share of `r round(els_jns_df %>% filter(oa_type == "Hybrid") %>% distinct(issn) %>% nrow() / length(unique(els_jns_df$issn)) * 100, 0)`%.

@@

Following the approach of our earlier study [@Matthias_2019], we started with APC pricing list information to determine the number of open access business model flips between 2015-2020. Although the transition from hybrid to a fully open access business model has drawn most attention, also the reverse can be observed. A total number of `r sum(flipped_jns_ind$n)` journals changed the business model under the ownership of Elsevier. Of these, `r filter(flipped_jns_ind, oa_model == "Open Access")$n` journals flipped to fully open access, while `r filter(flipped_jns_ind, oa_model != "Open Access")$n` changed to hybrid open access. Examples for such reverse flipped open access journals include the physical journal *Materials Today* and the life science journal *Meta Gene*.

@@

# oa
ind_per_year <- hybrid_df %>%
  filter(oa_archive == FALSE) %>%
  group_by(issued_year) %>%
  summarise(oa_articles = n()) %>%
  inner_join(els_yr, by = "issued_year") %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles * 100)
ind_summary_table <- ind_per_year %>% 
  pivot_longer(oa_articles:prop, names_to = "indicator", values_to = "count") %>%
  pivot_wider(names_from = issued_year, values_from = count) %>%
  mutate(indicator = factor(indicator, levels = c("journals", "articles", "oa_articles", "prop"))) %>%
  arrange(indicator) %>%
  mutate(indicator = recode(indicator,
                            oa_articles = "Hybrid OA Articles", 
                            articles = "All Articles", 
                            journals = "Hybrid journals with at least one OA article", 
                            prop = "Percentage"))
`Total` <- c(
  hybrid_df %>% filter(oa_archive == FALSE) %>% distinct(issn) %>% nrow(),
  sum(ind_summary_table[2,2:6]),
  sum(ind_summary_table[3,2:6]),
  sum(ind_per_year$oa_articles) / sum(ind_per_year$articles) * 100)
uptake_table <- bind_cols(ind_summary_table, Total = `Total`) %>%
  rename(` ` = indicator) %>%
  knitr::kable(digits = 1)
```

Overall, `r Total[1]` hybrid journals provided immediate open access to `r Total[3]` articles between 2015 and 2019. The total share of hybrid open access in the publication volume of Elsevier journals was `r Total[4]`%. Table present the high-level findings by year, illustrating a moderate growth in hybrid open access. Although the number of hybrid journals with at least one open access article remained somehow constant, the number of open access articles rose substantially from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_articles)` (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_articles)` (2019). However, the overall uptake of open access in Elsevier's hybrid journal portfolio grew slowly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(prop)`% (2015) to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(prop)`% (2019), because the number of articles published also increased during this five-years period.

```{r}
uptake_table
```

```{r}
els_jn_pub_volume <- els_pub_volume %>%
  filter(issn %in% eligible_jns$issn) %>%
  group_by(issued_year, issn) %>%
  summarise(articles = sum(articles),
            journals = n_distinct(issn))
hybrid_oa_per_jn <- hybrid_df %>%
  filter(oa_archive == FALSE) %>%
  group_by(issued_year, issn) %>%
  summarise(oa_articles = n())%>%
  inner_join(els_jn_pub_volume, by = c("issued_year", "issn")) %>%
  ungroup() %>%
  mutate(prop = oa_articles / articles * 100) %>% View()
```

The open access share varied across Elsevier hybrid journals. Since 2015, most hybrid journals have had a slow uptake rate of open access articles. In general, open access via the hybrid open access publishing model played a marginal role in the context of Elsevier’s total publication volume.

# license
